Forecasting the United States 2024 Presidential Election

### 2024-11-04 paper/paper.qmd
added section on clinton effect

>>>>

A further limitation of our model’s prediction is that there could be inherent biases present in the data which are difficult to locate due to the vast size of the compiled dataset. In this election, there is a greater demographic contrast between candidates than ever before, with a Black woman running against a White man. With such contrasting identities being represented, the respondents are more likely than ever to be affected by biases and prejudices based on the candidates’ race and gender. However, due to social desirability bias, many respondents may choose to not answer or answer more neutrally in order to conceal the strength of their disapproval for Harris due to these factors. This may cause a skew where predictions indicate significantly higher support for Harris than is actually present in the population.

Despite survey predictions successfully minimizing the effect of social desirability when predicting the election results in 2008 and 2012 for Barack Obama, it had a significant effect in 2016 when the only other woman, Hillary Clinton, ran for president [@clintoneffect]. The election that year saw that previously undecided voters largely leaned against Clinton, and finally ended with her loss [@clintoneffect]. With this in mind, it is uncertain how much of the neutral and non-responses within the current dataset are caused by this bias, meaning that the higher support this current model predicts for Harris also could be a misleading result.
